Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
s/fold/foldr/
cocosea

\newcommand{\theHalgorithm}{\arabic{algorithm}}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\system}{\textsc{ECC}~}
\newcommand{\systemEnding}{\textsc{ECC}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}

@@

Imagine you are asked  to edit some text, and
told that you should change the text ``Nancy FreeHafer'' to
``Dr. Nancy''.  From this example, you likely infer that ``Jane
Goodall'' goes to ``Dr. Jane'', drawing upon your prior knowledge of
text, like that words are separated by spaces. Few-shot learning problems
like these
are commonplace in both human and machine learning.
We cast these few-shot learning problems as each being a program synthesis problem.
The program synthesis and AI communities have built many successful program synthesis algorithms,
spanning text editing (e.g., FlashFill~\cite{gulwani2011automating}), motor programs~\cite{lake2015human}, procedural graphics~\cite{ellis2017learning}, planning
procedures~\cite{devlin2017neural}, and many others.  However, the
success of these systems hinges upon a carefully
hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
impart prior knowledge of a domain
by providing a restricted set
of finely-tuned programming primitives: for
%that are finely tuned to the problems in the domain: for
text editing,
these are primitives like appending and splitting on characters.
In this work,
we consider the problem of
building agents that solve program learning tasks,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
Our solution is an algorithm that learns a DSL
while jointly training a neural network to help write programs in the learned DSL.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Jane Goodall&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_1$\code{ s ' '))}\\\cmidrule{2-2}
    \sc{DSL}
&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y) (cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s ($\lambda$ (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a library of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}

%% We think of
%% solutions to these tasks as being well represented by programs,
%% and so our problem can be stated as follows:
%% how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good text editing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
We combine these two ideas into
an algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library (DSL) of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which DSL components are
useful for which kinds of tasks.
Together, the DSL and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
because it iterates between three different steps: an
\textbf{Explore} step uses the DSL and neural network to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} step
modifies the structure of the DSL by discovering regularities across programs found by
the previous Explore step; and a \textbf{Compile} step, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

Because any model may be encoded as a (deterministic or probabilistic) program,
we carefully delineate the scope of problems considered here.
We think of \system as learning to solve the kinds of problems that humans
can solve relatively quickly -- once they acquire the relevant domain expertise.
These correspond to short programs -- if you have an expressive DSL.
Even with a good DSL, program search may be intractable, so
we amortize the cost of program search by training a neural network to
assist the search procedure.

We apply \system to four domains:
list processing;
text editing (in the style of FlashFill~\cite{gulwani2011automating});
and symbolic regression.
For each of these we initially provide a generic set of programming primitives.
Our algorithm then discovers its own DSL for expressing solutions in the
domain (Tbl.~\ref{initialExampleDSL}).
%% \begin{figure}
%%   \begin{tabular}{lcl}
%%     \toprule
%%     Domain&Example Task&Part of the learned DSL\\\midrule
%%     Lists&
%%     \begin{tabular}{rcl}
%%       \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
%%       \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
%%     \end{tabular}
%%     &
%%     \begin{tabular}{l}
%%       \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
%%           \hspace{0.5cm} \emph{(appends lists)}
%%       \end{tabular}

%%     \\    Strings&
%%     \begin{tabular}{rcl}
%%       Temple Anna H& $\longrightarrow$& TAH\\
%%       Lara Gregori&$\longrightarrow$&LG
%%       \end{tabular}
%%     &
%%     \begin{tabular}{l}
%%       \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
%%       \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
%%       \end{tabular}

%%     \\Regression&\includegraphics[width = 1cm]{figures/functions/476.png}&\begin{tabular}{l}
%%       \code{(+ (* real x) real)} \\
%%           \hspace{0.5cm} \emph{(a linear function of x)}
%%       \end{tabular}

%%     \\\bottomrule
%%   \end{tabular}
%%   \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
%% %  \vspace{-4em}
%% \end{figure}

\begin{figure}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      % \code{[99\, 4\, 2]}$\to $\code{99}\\
      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[1]}$\to $\code{[1\, 3]}\\
      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

  \begin{tabular}{l}
    % $f_0($\code{r}$,\ell) \,=\, $\code{(fold $\ell$ r ($\lambda$ (x a)}\\
    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(fold}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
    % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
    %NOTE: this is the actual invention, but I removed a redundant lambda
    % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
    $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}\\
    \vspace{2ex}
      \hspace{0.5cm}($f_0$: \emph{Concatenate }\code{r}\emph{ with $\ell$})\\
    $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
    \vspace{2ex}
      \hspace{0.5cm}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
    $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
    \phantom{$f_1(\ell,$\code{k}}
    \code{($\lambda$ (x a) (if a a (= k x))))}\\
      \hspace{0.5cm}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
  \end{tabular}&

@@

\end{tabular}}
\caption{Top: Example tasks from each domain, each followed by the programs
  \system discovers for them. Bottom: Subset of learned DSL. Notice that
  learned DSL primitives can call each other
  ($f_0$/$f_1$ in Lists, $f_1$/$f_2$ in Text \& Regression)}
\end{figure}

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We were inspired by EC,
and go beyond it by giving a new algorithm for learning DSLs, as well as a way of combining DSL learning with
neurally guided program search.
In the experiments section,
we compared directly with EC,
showing empirical improvements.
The contribution of this work is
then
an effective algorithm for learning DSLs which jointly trains a neural net to
search for programs in the DSL.

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration primarily from the
Exploration-Compression algorithm for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We extend these ideas into an inference strategy that iterates
through three steps: an \textbf{Explore} step uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Compress} and \textbf{Compile} steps update the DSL and the
recognition model, respectively.  Crucially, these steps bootstrap 
each other:
\begin{comment}
  \begin{wrapfigure}{r}{0.5\textwidth}
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

\begin{wrapfigure}{R}{0.6\textwidth}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %($\lambda$ ($\lambda$ (fold $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some text editing learned subroutines }\label{textPrimitives}
  \end{table}

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
%% For symbolic regression, we also care about the quality
%% of the solution, as measured by the likelihood model $\probability[x|p]$,
%% e.g. did the agent correctly explain a linear function
%% using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these baselines.
Our full model consistently
improves on the baselines,
sometimes dramatically (text editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,
taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports a view of the recognition model as a way of amortizing the cost of searching for programs.

>>>>

\newcommand{\theHalgorithm}{\arabic{algorithm}}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\system}{\textsc{CoCoSea}~}
\newcommand{\systemEnding}{\textsc{CoCoSea}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}

@@

Imagine you are asked  to edit some text, and
told that you should change the text ``Nancy FreeHafer'' to
``Dr. Nancy''.  From this example, you likely infer that ``Jane
Goodall'' goes to ``Dr. Jane'', drawing upon your prior knowledge of
text, like that words are separated by spaces. Few-shot learning problems
like these
are commonplace in both human and machine learning.
We cast these few-shot learning problems as each being a program synthesis problem.
The programming languages and AI communities have built many successful program synthesis algorithms,
spanning text editing (e.g., FlashFill~\cite{gulwani2011automating}), motor programs~\cite{lake2015human}, procedural graphics~\cite{ellis2017learning}, planning
procedures~\cite{devlin2017neural}, and many others.  However, the
success of these systems hinges upon a carefully
hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
impart prior knowledge of a domain
by providing a restricted set
of finely-tuned programming primitives: for
%that are finely tuned to the problems in the domain: for
text editing,
these are primitives like appending and splitting on characters.
In this work,
we consider the problem of
building agents that solve program learning tasks,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
Our solution is an algorithm that learns a DSL
while jointly training a neural network to help write programs in the learned DSL.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Jane Goodall&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_1$\code{ s ' '))}\\\cmidrule{2-2}
    \sc{DSL}
&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b ($\lambda$ (x y) (cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a library of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}

%% We think of
%% solutions to these tasks as being well represented by programs,
%% and so our problem can be stated as follows:
%% how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good text editing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
We weave these ideas together into
an algorithm called \system (\textbf{Co}mpress/\textbf{Co}mpile/\textbf{Sea}rch),
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Searching for programs that solve the tasks;
(2) Composing a library (DSL) of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which DSL components are
useful for which kinds of tasks.
Together, the DSL and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

%% Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
%% because it
\system iterates through three different steps: a
\textbf{Search} step uses the DSL and neural network to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} step
modifies the structure of the DSL by discovering regularities across programs found during search; and a \textbf{Compile} step, which
improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related programming problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.
%% Because any model may be encoded as a (deterministic or probabilistic) program,
%% we carefully delineate the scope of problems considered here.
We think of \system as learning to solve the kinds of problems that humans
can solve relatively quickly -- once they acquire the relevant domain expertise.
These correspond to short programs -- if you have an expressive DSL.
Even with a good DSL, program search may be intractable, so
we amortize the cost of search by training a neural network to
assist the search procedure.

We apply \system to three domains:
list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then constructs
its own DSL for expressing solutions in the domain (Fig.~\ref{initialExampleDSL}).
\begin{figure}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      % \code{[99\, 4\, 2]}$\to $\code{99}\\
      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[1]}$\to $\code{[1\, 3]}\\
      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

  \begin{tabular}{l}
    % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
    % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
    %NOTE: this is the actual invention, but I removed a redundant lambda
    % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
    $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}\\
    \vspace{2ex}
      \hspace{0.5cm}($f_0$: \emph{Concatenate }\code{r}\emph{ with $\ell$})\\
    $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
    \vspace{2ex}
      \hspace{0.5cm}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
    $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
    \phantom{$f_1(\ell,$\code{k}}
    \code{($\lambda$ (x a) (if a a (= k x))))}\\
      \hspace{0.5cm}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
  \end{tabular}&

@@

\end{tabular}}
\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them. Bottom: Subset of learned DSL. Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
\end{figure}

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We were inspired by EC,
and go beyond it by giving a new algorithm that learns DSLs, as well as a way of combining DSL learning with
neurally guided program search.
In the experiments section,
we compared directly with EC,
showing empirical improvements.

The new contribution of this work is
thus an  algorithm for learning DSLs which jointly trains a neural net to
search for programs in the DSL.

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration primarily from the
Exploration-Compression algorithm for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We extend these ideas into an inference strategy that iterates
through three steps: a \textbf{Search} step uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Compress} and \textbf{Compile} steps update the DSL and the
recognition model, respectively.  Crucially, these steps bootstrap 
each other:
\begin{comment}
  \begin{wrapfigure}{r}{0.5\textwidth}
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

\begin{wrapfigure}{R}{0.6\textwidth}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Search})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Search)}}

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %($\lambda$ ($\lambda$ (foldr $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some text editing learned subroutines }\label{textPrimitives}
  \end{table}

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
%% For symbolic regression, we also care about the quality
%% of the solution, as measured by the likelihood model $\probability[x|p]$,
%% e.g. did the agent correctly explain a linear function
%% using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these baselines.
Our full model consistently
improves on the baselines,
sometimes dramatically (text editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,

taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports a view of the recognition model as a way of amortizing the cost of searching for programs.
